---
title: "COVIDCave28OctRevision"
author: "Adam Dennett"
date: "28/10/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(geojsonio)
library(tmap)
library(downloader)
library(readxl)
library(janitor)
```

# New COVID Case Data Maps

## Get some COVID-19 MSOA-level Case and Boundary Data:

```{r}
COVID_MSOA_Tidy <- read_csv("https://coronavirus.data.gov.uk/downloads/msoa_data/MSOAs_latest.csv")

#get some boundaries
#super-generalised url to geojson
url <- "https://opendata.arcgis.com/datasets/87aa4eb6393644768a5f85929cc704c2_0.geojson"

#read those boundaries into an sf() object for plotting
UK_MSOA_sf <- geojson_sf(url)
```

Join those boundaries and data together

```{r mapJoin}
#join the MSOA boundaries to the data
COVID_MSOA_Tidy_Map <- UK_MSOA %>% 
  merge(.,
        COVID_MSOA_Tidy,
        by.x="MSOA11CD",
        by.y="areaCode")
```

## First Test Map

```{r quickMap, echo=FALSE}
#make a very quick and dirty interactive map of week 41
tmap_mode("view")
COVID_MSOA_Tidy_Map %>% filter(date == "2020-10-24") %>% 
tm_shape() +
  tm_fill(col = "newCasesBySpecimenDateRollingRate", style="fisher")
```

## MSOA Rates and Lockdown Tiers

Get some lockdown tier data

```{r tierData}

la_lockdown_tiers <- read_csv("https://visual.parliament.uk/research/visualisations/coronavirus-restrictions-map/commonslibrary-coronavirus-restrictions-data.csv")

spec(la_lockdown_tiers)
```

Pull out the national data from here:

```{r pullOutNationalData}

national_lockdown <- la_lockdown_tiers %>% 
  filter(l_restrictions == "National")
```

Get some Local Authority Boundary Data

```{r}


#read those boundaries into an sf() object for plotting
#not generalised that much
UK_lad_sf <- geojson_sf("https://opendata.arcgis.com/datasets/3b374840ce1b4160b85b8146b610cd0c_0.geojson")
#generalised quite a lot
#UK_lad_sf <- geojson_sf("https://opendata.arcgis.com/datasets/910f48f3c4b3400aa9eb0af9f8989bbe_0.geojson")

```

Merge it with the LAD boundaries

```{r}
LAD_COVID_Restriction_Map <- left_join(UK_lad_sf, la_lockdown_tiers, by = c("LAD20CD"= "l_l_lockdownareacode"), keep=T)
```

OK, now we have loads of empty rows, replace the local values with national values if empty - please fix this code for me!

```{r}
temp_1 <- LAD_COVID_Restriction_Map %>% 
  filter(is.na(l_l_widerlacode))
temp_2 <- LAD_COVID_Restriction_Map %>% 
  filter(!is.na(l_l_widerlacode))

#names(temp_1)
#names(national_lockdown)

temp_1[str_detect(temp_1$LAD20CD,"^E"),11:40] <- national_lockdown[which(!is.na(str_match(national_lockdown$l_Country, "E"))),1:30]
temp_1E <- temp_1[str_detect(temp_1$LAD20CD,"^E"),]
temp_1W <- temp_1[str_detect(temp_1$LAD20CD,"^W"),]
temp_1S <- temp_1[str_detect(temp_1$LAD20CD,"^S"),]
temp_1N <- temp_1[str_detect(temp_1$LAD20CD,"^N"),]

temp_1E[,11:40] <- national_lockdown[which(!is.na(str_match(national_lockdown$l_Country, "E"))),1:30]
temp_1W[,11:40] <- national_lockdown[which(!is.na(str_match(national_lockdown$l_Country, "W"))),1:30]
temp_1S[,11:40] <- national_lockdown[which(!is.na(str_match(national_lockdown$l_Country, "S"))),1:30]
temp_1N[,11:40] <- national_lockdown[which(!is.na(str_match(national_lockdown$l_Country, "N"))),1:30]

LAD_COVID_Restriction_Map1 <- rbind(temp_1E,temp_1W,temp_1S,temp_1N,temp_2)

#find the row where the value matching E is located in the l_country column
#which(!is.na(str_match(national_lockdown$l_Country, "E")))
#ifelse(str_detect(temp_1$LAD20CD,"^E"),national_lockdown[which(!is.na(str_match(national_lockdown$l_Country, "E"))),1:30],"nope")

#names(national_lockdown)
```

LAD Lockdown Map + MSOA-level rolling case counts

```{r}
tmap_mode("view")

tm_shape(LAD_COVID_Restriction_Map1) +
  tm_fill(col = "l_tier", style = "cat", palette=c("#fd8d3c","#ffffb2","#bd0026"), alpha = 0.7, popup.vars = "LAD20NM") + 
tm_shape(COVID_MSOA_Tidy_Map) +
  tm_dots(size = "newCasesBySpecimenDateRollingSum", alpha = 0.5,
  border.alpha = 0, col = "#bae4b3", clustering = T)
  
```

## And we can make some faceted maps as well

```{r}
#now plot a faceted map using ggplot facet map
ggplot(data = COVID_MSOA_Tidy_Map,aes(fill=newCasesBySpecimenDateRollingSum)) +
  geom_sf(color='transparent') +
  scale_fill_viridis_c(option = 'plasma')+
  facet_wrap(~date)

```

```{r}
#how about if we rank the data like Rich Harris did?
range(COVID_MSOA_Tidy_Map$newCasesBySpecimenDateRollingRate)

#create a new rank column
COVID_MSOA_Tidy_Map$rate_rank <- rank(COVID_MSOA_Tidy_Map$newCasesBySpecimenDateRollingRate)

#now plot a ranked faceted map using ggplot facet map
#lets go for 4 rows this time like Rich did...
ggplot(data = COVID_MSOA_Tidy_Map,aes(fill=rate_rank)) +
  geom_sf(color='transparent') +
  scale_fill_viridis_c(option = 'plasma')+
  facet_wrap(~date, nrow = 4)
```

