---
title: "Covid Cave"
author: "Adam Dennett"
date: "26/10/2020"
output: html_document
---

# Dr D's COVID Cave

## Creating Some Faceted Maps

### Setting Everything Up

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(geojsonio)
library(tmap)
```

```{r dataSetup}
#load covid data from the UK Government Website
COVID_MSOA <- read_csv("https://coronavirus.data.gov.uk/downloads/msoa_data/MSOAs_latest.csv")

#*Data Cleaning Task
#*
#*Think about how we could useful re-code -99 
#*which currently codes cases between 0-2
#*it might be that a new re-coded column is created in the
#*long tidy version of the data

#get some boundaries
#super-generalised url to geojson
url <- "https://opendata.arcgis.com/datasets/87aa4eb6393644768a5f85929cc704c2_0.geojson"

#read those boundaries into an sf() object for plotting
UK_MSOA <- geojson_sf(url)
```

### Joining boundaries and data together

```{r mapJoin}
#join the MSOA boundaries to the data
MSOA_COVID_Map <- UK_MSOA %>% 
  merge(.,
        COVID_MSOA,
        by.x="MSOA11CD",
        by.y="msoa11_cd")

```

## First Test Map

```{r quick map, echo=FALSE}
#make a very quick and dirty interactive map of week 41
tmap_mode("view")
tm_shape(MSOA_COVID_Map) +
  tm_fill(col = "wk_41", style="jenks")
```

An alternative with ggplot2

```{r ggplotMap}
#here's a quick and dirty map with a continuous scale
#using ggplot2
ggplot(data = MSOA_COVID_Map,aes(fill=wk_41)) +
  geom_sf(color='transparent') +
  scale_fill_viridis_c(option = 'plasma')
```

## Creating Some Faceted Maps

Check the data looks OK first

```{r CheckDataTypes}
types <- sapply(MSOA_COVID_Map, class)
```

The data we want to plot starts from column 9 onwards therefore, pivot from column 9 onwards:

```{r pivotLong}
MSOA_COVID_Long <- COVID_MSOA %>% 
  pivot_longer(.,
               c(9:46))
```

Now re-merge the boundary geometries to the long file:

```{r reMerge}
MSOA_COVID_Map_Long <- UK_MSOA %>% 
  merge(.,
        MSOA_COVID_Long,
        by.x="MSOA11CD",
        by.y="msoa11_cd")
```

Now plot a faceted map using ggplot facet map

```{r facetMap}
ggplot(data = MSOA_COVID_Map_Long,aes(fill=value)) +
  geom_sf(color='transparent') +
  scale_fill_viridis_c(option = 'plasma')+
  facet_wrap(~name)
```

We could try ranging the data like Rich Harris did:

```{r}
range(MSOA_COVID_Map_Long$value)
#create a new rank column
MSOA_COVID_Map_Long$value_rank <- rank(MSOA_COVID_Map_Long$value)
```

Now plot a ranked faceted map using ggplot facet map and lets go for 4 rows this time like Rich did...

```{r}
ggplot(data = MSOA_COVID_Map_Long,aes(fill=value_rank)) +
  geom_sf(color='transparent') +
  scale_fill_viridis_c(option = 'plasma')+
  facet_wrap(~name, nrow = 4)

#get some boundaries
#super-generalised url to geojson
url <- "https://opendata.arcgis.com/datasets/87aa4eb6393644768a5f85929cc704c2_0.geojson"

#read those boundaries into an sf() object for plotting
UK_MSOA <- geojson_sf(url)
```

### Joining boundaries and data together

```{r map join}
#join the MSOA boundaries to the data
MSOA_COVID_Map <- UK_MSOA %>% 
  merge(.,
        COVID_MSOA,
        by.x="MSOA11CD",
        by.y="msoa11_cd")

```

## First Test Map

```{r quick map, echo=FALSE}
#make a very quick and dirty interactive map of week 41
tmap_mode("view")
tm_shape(MSOA_COVID_Map) +
  tm_fill(col = "wk_41", style="jenks")
```

An alternative with ggplot2

```{r}
#here's a quick and dirty map with a continuous scale
#using ggplot2
ggplot(data = MSOA_COVID_Map,aes(fill=wk_41)) +
  geom_sf(color='transparent') +
  scale_fill_viridis_c(option = 'plasma')
```
