---
title: "Covid Cave - Week 4 Project"
author: "Isabella Brant"
date: "28/10/2020"
output: html_document
---

# Dr D's COVID Cave

## Creating Some Faceted Maps

### Setting Everything Up

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(geojsonio)
library(tmap)
```

```{r dataSetup}

#load covid data from the UK Government Website
COVID_MSOA <- read_csv("https://coronavirus.data.gov.uk/downloads/msoa_data/MSOAs_latest.csv")

## Creating a covid_msoa dataset for the week 2020-10-23
COVID_MSOA_20201023 <- COVID_MSOA %>%
  filter(str_detect(`date`, "2020-10-23"))

#*Data Cleaning Task
#*
#*Think about how we could useful re-code -99 
#*which currently codes cases between 0-2
#*it might be that a new re-coded column is created in the
#*long tidy version of the data
#get some boundaries
#super-generalised url to geojson
url <- "https://opendata.arcgis.com/datasets/87aa4eb6393644768a5f85929cc704c2_0.geojson"
#read those boundaries into an sf() object for plotting
UK_MSOA <- geojson_sf(url)

```

### Joining boundaries and data together

```{r mapJoin}

#join the MSOA boundaries to the data
MSOA_COVID_Map <- UK_MSOA %>% 
  merge(.,
        COVID_MSOA,
        by.x="MSOA11CD",
        by.y="areaCode")


#join the MSOA boundaries to the 2020-10-23 data
MSOA_COVID_Map_20201023 <- UK_MSOA %>% 
  merge(.,
        COVID_MSOA_20201023,
        by.x="MSOA11CD",
        by.y="areaCode")

```

## First Test Map

```{r quickMap, echo=FALSE}

#make a very quick and dirty interactive map of week/c 2020-10-23
tmap_mode("view")
tm_shape(MSOA_COVID_Map_20201023) +
  tm_fill(col = "newCasesBySpecimenDateRollingSum", style="jenks")

```

## Calculating COVID Rates

In order to calculate rates we need to determine a 'population at risk' of contracting COVID-19. For this disease, the population at risk is everyone. This isn't always the case - for example, if we were looking at dementia, we wouldn't include children in the population at risk.

### Download population at risk data

The Office for National Statistics create mid-year population estimates. We will download these in order to calculate our rates:

```{r downloadPopData}

library(downloader)
library(readxl)
library(janitor)
url_msoa_pop_zip <- "https://www.ons.gov.uk/file?uri=%2fpeoplepopulationandcommunity%2fpopulationandmigration%2fpopulationestimates%2fdatasets%2fmiddlesuperoutputareamidyearpopulationestimates%2fmid2019sape22dt4/sape22dt4mid2019msoasyoaestimatesunformatted.zip"
#download from the URL
download(url_msoa_pop_zip, dest="dataset.zip", mode="wb")
#unzip into a new directory called data
unzip("dataset.zip",exdir="./data")
#get the file names from within the zip file
filename <- list.files("./data")
#read the sheet you want from the file
msoa_pop <- read_excel(file.path("./data",filename), sheet="Mid-2019 Persons", skip = 3, col_names = T)
msoa_pop <- clean_names(msoa_pop)

```

Now we can join this population at risk data to our COVID-19 case data. The following code chunk selects just the 'all_ages' column from the msoa_pop population data and joins it using left join to the pivoted 'tidy' long version of the COVID-19 case data:

```{r mergePopData}

COVID_MSOA <- left_join(COVID_MSOA,
                              msoa_pop %>% select(msoa_code,all_ages),
                              by = c("areaCode" = "msoa_code"))


## Creating a new column for covid rate per 100,000
COVID_MSOA <- COVID_MSOA%>%
  mutate(case_rate = (newCasesBySpecimenDateRollingSum / all_ages) * 100000)


## Joining the updated COVID_MSOA data with UK_MSOA boundary data
MSOA_COVID_Map_Case_Rate <- UK_MSOA %>% 
  merge(.,
        COVID_MSOA,
        by.x="MSOA11CD",
        by.y="areaCode")


## Subsetting the COVID_MSOA for w/c 2020-10-23 inc case_rate
COVID_MSOA_20201023_Case_Rate <- COVID_MSOA%>%
  filter(str_detect(`date`, "2020-10-23"))


## Joining the COVID_MSOA for w/c 2020-10-23 inc case_rate with UK_MSOA boundary data
MSOA_COVID_Map_Case_Rate_20201023 <- UK_MSOA %>% 
  merge(.,
        COVID_MSOA_20201023_Case_Rate,
        by.x="MSOA11CD",
        by.y="areaCode")


## Creating a map for case_rate per 100,000
tmap_mode("view")
tm_shape(MSOA_COVID_Map_Case_Rate_20201023) +
  tm_fill(col = "case_rate", style="fisher")


```
