---
title: "Covid Cave"
author: "Adam Dennett"
date: "26/10/2020"
output: html_document
---

# Dr D's COVID Cave

## Creating Some Faceted Maps

### Setting Everything Up

------------------------------------------------------------------------

*Warning - this code has been superseded on 28th Oct as the government have now decided to release their weekly data in a new Tidy format!*

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(geojsonio)
library(tmap)
```

```{r dataSetup}
#load covid data from the UK Government Website
COVID_MSOA <- read_csv("https://coronavirus.data.gov.uk/downloads/msoa_data/MSOAs_latest.csv")

#*Data Cleaning Task
#*
#*Think about how we could useful re-code -99 
#*which currently codes cases between 0-2
#*it might be that a new re-coded column is created in the
#*long tidy version of the data

#get some boundaries
#super-generalised url to geojson
url <- "https://opendata.arcgis.com/datasets/87aa4eb6393644768a5f85929cc704c2_0.geojson"

#read those boundaries into an sf() object for plotting
UK_MSOA_sf <- geojson_sf(url)
```

### Joining boundaries and data together

```{r mapJoin}
#join the MSOA boundaries to the data
MSOA_COVID_Map <- UK_MSOA_sf %>% 
  merge(.,
        COVID_MSOA,
        by.x="MSOA11CD",
        by.y="msoa11_cd")

```

## First Test Map

```{r quickMap, echo=FALSE}
#make a very quick and dirty interactive map of week 41
tmap_mode("view")
tm_shape(MSOA_COVID_Map) +
  tm_fill(col = "wk_41", style="jenks")
```

An alternative with ggplot2

```{r ggplotMap}
#here's a quick and dirty map with a continuous scale
#using ggplot2
ggplot(data = MSOA_COVID_Map,aes(fill=wk_41)) +
  geom_sf(color='transparent') +
  scale_fill_viridis_c(option = 'plasma')
```

## Creating Some Faceted Maps

Check the data looks OK first

```{r CheckDataTypes}
types <- sapply(MSOA_COVID_Map, class)
```

The data we want to plot starts from column 9 onwards therefore, pivot from column 9 onwards:

```{r pivotLong}
MSOA_COVID_Long <- COVID_MSOA %>% 
  pivot_longer(.,
               c(9:length(COVID_MSOA)))

```

Now re-merge the boundary geometries to the long file:

```{r reMerge}
MSOA_COVID_Map_Long <- UK_MSOA_sf %>% 
  merge(.,
        MSOA_COVID_Long,
        by.x="MSOA11CD",
        by.y="msoa11_cd")
```

Now plot a faceted map using ggplot facet map

```{r facetMap}
ggplot(data = MSOA_COVID_Map_Long,aes(fill=value)) +
  geom_sf(color='transparent') +
  scale_fill_viridis_c(option = 'plasma')+
  facet_wrap(~name)
```

We could try ranging the data like Rich Harris did:

```{r}
range(MSOA_COVID_Map_Long$value)
#create a new rank column
MSOA_COVID_Map_Long$value_rank <- rank(MSOA_COVID_Map_Long$value)
```

Now plot a ranked faceted map using ggplot facet map and lets go for 4 rows this time like Rich did...

```{r}
ggplot(data = MSOA_COVID_Map_Long,aes(fill=value_rank)) +
  geom_sf(color='transparent') +
  scale_fill_viridis_c(option = 'plasma')+
  facet_wrap(~name, nrow = 4)
```

## Calculating COVID Rates

In order to calculate rates we need to determine a 'population at risk' of contracting COVID-19. For this disease, the population at risk is everyone. This isn't always the case - for example, if we were looking at dementia, we wouldn't include children in the population at risk.

### Download population at risk data

The Office for National Statistics create mid-year population estimates. We will download these in order to calculate our rates:

```{r downloadPopData}
library(downloader)
library(readxl)
library(janitor)

url_msoa_pop_zip <- "https://www.ons.gov.uk/file?uri=%2fpeoplepopulationandcommunity%2fpopulationandmigration%2fpopulationestimates%2fdatasets%2fmiddlesuperoutputareamidyearpopulationestimates%2fmid2019sape22dt4/sape22dt4mid2019msoasyoaestimatesunformatted.zip"

#download from the URL
download(url_msoa_pop_zip, dest="dataset.zip", mode="wb")

#unzip into a new directory called data
unzip("dataset.zip",exdir="./data")

#get the file names from within the zip file
filename <- list.files("./data")

#read the sheet you want from the file
msoa_pop <- read_excel(file.path("./data",filename), sheet="Mid-2019 Persons", skip = 3, col_names = T)

msoa_pop <- clean_names(msoa_pop)

```

Now we can join this population at risk data to our COVID-19 case data. The following code chunk selects just the 'all_ages' column from the msoa_pop population data and joins it using left join to the pivoted 'tidy' long version of the COVID-19 case data:

```{r mergePopData}

MSOA_COVID_Long <- left_join(MSOA_COVID_Long,
                              msoa_pop %>% select(msoa_code,all_ages),
                              by = c("msoa11_cd" = "msoa_code"))
  
```

Replace the -99 values with missing data

```{r replaceValues}
MSOA_COVID_Long <- MSOA_COVID_Long %>% 
  mutate(value=na_if(value,-99))

```

Now create a new rate column:

```{r covidRate}

MSOA_COVID_Long <- MSOA_COVID_Long %>% 
  mutate(case_rate = (value / all_ages)*100000)

```

```{r reMerge}
MSOA_COVID_Map_Long <- UK_MSOA_sf %>% 
  merge(.,
        MSOA_COVID_Long,
        by.x="MSOA11CD",
        by.y="msoa11_cd")
```

Make a map of the last 7 days

```{r quickMap, echo=FALSE}
#make a very quick and dirty interactive map of week 41
tmap_mode("view")
MSOA_COVID_Map_Long %>% filter(name == "latest_7_days") %>% 
tm_shape() +
  tm_fill(col = "case_rate", style="fisher")
```
